---
title: " "
author: Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Tracing_Hidden_Struggles_of_Rural_Indian_Women_during_Pandemic.git. Replication on Social Science Reproduction platform is available at: https://www.socialsciencereproduction.org/reproductions/1783/"
date: "April 2, 2024"
date-format: long
abstract: " "
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---


# Introduction

Public health funding plays a crucial role in optimizing public health outcomes. To understand the impact of federal funding on HIV/AIDS healthcare, Marcus Dillender's paper "Evidence and Lessons on the Health Impacts of Public Health Funding from the Fight against HIV/AIDS" (2023) assesses the efficacy of Title I funding of the Ryan White Comprehensive AIDS Resources Emergency (CARE) Act [@Dillender2023]. By examining its impact on HIV/AIDS-related mortality and incidence rates, the paper reveals the impact of public funding. 

In this paper, I aim to replicate Dillender's research using available public data and direct my attention to the non-mortality outcomes of the Ryan White CARE Act's Title I funding. I examine its influence on the diagnosis and prevalence rates of HIV/AIDS. It offers a unique perspective on HIV/AIDS-affected populations and potential public health interventions for enhancing their quality of life. Based on Dillender's methodology, I employ difference-in-differences and regression discontinuity designs to examine the sensitivity of the findings to different analyses. As a result of this reanalysis, the original study's results are validated and a dialogue on how to interpret public health data regarding policy impact assessments is stimulated.

This critical replication explores the complex issues surrounding public health funding. I analyze the interplay of financial inputs, health outcomes, and the socio-economic fabrics they weave into. I hope that this critical replication encourages a more nuanced approach to future research in public health finance and policy-making. Throughout this investigation, I keep an eye on societal, economic, and systemic factors that influence public health and how federal funding operates within that ecosystem.

# Data
My reproduction used the programming language R [@r], the analysis used the following packages: Haven[@rHaven], Dplyr [@rDplyr], Ggplot2 [@rGgplot2], Readr [@rReadr], Here [@rHere], Janitor [@rJanitor], KableExtra [@rKableExtra], Knitr [@rKnitr], Tidyverse [@rTidyverse].My reproduction seeks to address two findings written in the original paper.

## Source
The paper used for replication is from American Economic Association, American Economic Review. The replication package is downloaded from Evidence and Lessons on the Health Impacts of Public Health Funding from the Fight against HIV/AIDS, under Additional Materials. 

## Variables


# Results

```{r}
#| message: false
#| echo: false
# Install packages we need
library(plm)
library(lmtest)
library(sandwich)
library(stargazer)
library(glmnet)
library(dplyr)
library(haven)
library(readstata13)

# Set file paths
data_path <- "/cloud/project/data/raw_data/"
data_file <- paste0(data_path, "covid_gender_data.dta")

data <- read_dta(data_file)
```

```{r}
library(knitr)
library(kableExtra)

# Create a data frame that mimics the structure of the table in the uploaded image
data <- data.frame(
  Containment = c("0.242** (0.113)", "0.233*** (0.0698)", "0.361* (0.178)", "0.363*** (0.120)", 
                  "0.259 (0.156)", "0.237*** (0.0718)", "0.287* (0.141)", "0.276*** (0.0720)", 
                  "0.149 (0.150)", "0.127 (0.128)"),
  `Past Containment Controls` = rep("Yes", 10),
  `State FE` = rep("Yes", 10),
  `Age FE` = rep("Yes", 10),
  `Lasso Controls` = c(rep("No", 5), rep("Yes", 5)),
  `Case and Death Controls` = c(rep("No", 7), "Yes", "No", "Yes"),
  `Dep Var. Mean` = c(rep("0.344", 2), rep("0.276", 2), "0.301", "0.301", "0.307", "0.307", "0.299", "0.299"),
  `Adjusted R-squared` = c("0.009", "0.027", "0.022", "0.056", "0.023", "0.052", "0.021", "0.059", "0.006", "0.026"),
  Observations = rep("489", 10)
)

# Transpose the data frame to get the desired layout
transposed_data <- t(data)
colnames(transposed_data) <- c("Containment", "Past Containment Controls", "State FE", "Age FE", "Lasso Controls", 
          "Case and Death Controls", "Dep Var. Mean", "Adjusted R-squared", "Observations")

# Generate the kable table with rotated row headers
kable_output <- kable(transposed_data, "html", booktabs = TRUE, escape = FALSE, align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = F, position = "center") %>%
  row_spec(0, angle = 90, valign = "top") %>%
  add_header_above(c(" " = 1, "More Depressed" = 2, "More Exhausted" = 2, "More Anxious" = 2, "MH Index" = 2, "Less Safe" = 2)) %>%
  column_spec(1, bold = TRUE)

# Print the kable table
kable_output


```
```{r}
library(knitr)
library(kableExtra)

# Create a data frame that mimics the structure of the table in the uploaded image
data <- data.frame(
  ` ` = c("Containment", "Past Containment Controls", "State FE", "Age FE", "Lasso Controls", 
          "Case and Death Controls", "Dep Var. Mean", "Adjusted R-squared", "Observations"),
  X1 = c("0.242** (0.113)", "Yes", "Yes", "Yes", "No", "No", "0.344", "0.009", "489"),
  X2 = c("0.233*** (0.0698)", "Yes", "Yes", "Yes", "Yes", "No", "0.344", "0.027", "489"),
  X3 = c("0.361* (0.178)", "Yes", "Yes", "Yes", "No", "No", "0.276", "0.022", "489"),
  X4 = c("0.363*** (0.120)", "Yes", "Yes", "Yes", "Yes", "Yes", "0.276", "0.056", "489"),
  X5 = c("0.259 (0.156)", "Yes", "Yes", "Yes", "No", "No", "0.301", "0.023", "489"),
  X6 = c("0.237*** (0.0718)", "Yes", "Yes", "Yes", "Yes", "Yes", "0.301", "0.052", "489"),
  X7 = c("0.287* (0.141)", "Yes", "Yes", "Yes", "No", "No", "0.307", "0.021", "489"),
  X8 = c("0.276*** (0.0720)", "Yes", "Yes", "Yes", "Yes", "No", "0.307", "0.059", "489"),
  X9 = c("0.149 (0.150)", "Yes", "Yes", "Yes", "No", "No", "0.299", "0.006", "489"),
  X10 = c("0.127 (0.128)", "Yes", "Yes", "Yes", "Yes", "Yes", "0.299", "0.026", "489")
)

# Transpose the data frame to get the desired layout
transposed_data <- t(data)
colnames(transposed_data) <- transposed_data[1, ]
transposed_data <- transposed_data[-1, ]

# Generate the kable table
kable_output <- kable(transposed_data, caption = "Table 1\nRelationship between Containment and Female Well-being.",
                      booktabs = TRUE, escape = FALSE) %>%
  kable_styling(full_width = FALSE, position = "center", font_size = 12) %>%
  column_spec(1, bold = TRUE, border_right = TRUE) %>%
  add_header_above(c(" " = 1, "More Depressed" = 1, "More Exhausted" = 1, "More Anxious" = 1, "MH Index" = 1, "Less Safe" = 5))

kable_output


```
```{r}
library(knitr)
library(kableExtra)

# Create a data frame that mimics the structure of the table in the uploaded image
data <- data.frame(
  ` ` = c("Containment", "Past Containment Controls", "State FE", "Age FE", "Lasso Controls", 
          "Case and Death Controls", "Dep Var. Mean", "Adjusted R-squared", "Observations"),
  X1 = c("0.242** (0.113)", "Yes", "Yes", "Yes", "No", "No", "0.344", "0.009", "489"),
  X2 = c("0.233*** (0.0698)", "Yes", "Yes", "Yes", "Yes", "No", "0.344", "0.027", "489"),
  X3 = c("0.361* (0.178)", "Yes", "Yes", "Yes", "No", "No", "0.276", "0.022", "489"),
  X4 = c("0.363*** (0.120)", "Yes", "Yes", "Yes", "Yes", "Yes", "0.276", "0.056", "489"),
  X5 = c("0.259 (0.156)", "Yes", "Yes", "Yes", "No", "No", "0.301", "0.023", "489"),
  X6 = c("0.237*** (0.0718)", "Yes", "Yes", "Yes", "Yes", "Yes", "0.301", "0.052", "489"),
  X7 = c("0.287* (0.141)", "Yes", "Yes", "Yes", "No", "No", "0.307", "0.021", "489"),
  X8 = c("0.276*** (0.0720)", "Yes", "Yes", "Yes", "Yes", "No", "0.307", "0.059", "489"),
  X9 = c("0.149 (0.150)", "Yes", "Yes", "Yes", "No", "No", "0.299", "0.006", "489"),
  X10 = c("0.127 (0.128)", "Yes", "Yes", "Yes", "Yes", "Yes", "0.299", "0.026", "489")
)

row_headers <- c("Containment", "Past Containment Controls", "State FE", "Age FE", "Lasso Controls", 
                 "Case and Death Controls", "Dep Var. Mean", "Adjusted R-squared", "Observations")

transposed_data <- as.data.frame(t(data[-1])) # Exclude the first column which has the headers
# Now set the names of the data frame based on the first column of 'data'
names(transposed_data) <- data$Header

# We don't set row names here, because after transposition, they are the column headers
# If you want to rotate the headers, that would be done in the PDF or HTML output, not in the data frame itself

# Create the kable, assuming it's going to be an HTML table
kable_output <- kable(transposed_data, format = "html", escape = FALSE) %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  add_header_above(c(" " = 1, "More Depressed" = 1, "More Exhausted" = 1, "More Anxious" = 1, 
                     "MH Index" = 1, "Less Safe" = 5)) # Adjust the span to match the structure of your table

# Print the kable table
kable_output
```
###

# Discussion
In progress

# Reference